% this file is called up by thesis.tex
% content in this file will be fed into the main document

%: ----------------------- name of chapter  -------------------------
\chapter{Conclusions}\label{conclusioni} % top level followed by section, subsection




%: ----------------------- contents from here ------------------------
%In this thesis, we evaluated the performance of M2MShare, a protocol which implements DTN techniques in mobile phones world, in order to enable a peer-to-peer file sharing system between peers in the network making leverage on their mobility and opportunistic contacts among them. To do so we implemented the protocol in the Oppostunistic Network Environment (ONE) simulator. This simulator is able to emulate human movements adopting several movement models in a map-based environment and allowing us to evaluate M2MShare behaviour in a realistic city-scale scenario.
In this thesis, we evaluated the performance of M2MShare, a DTN solution for mobile phones that leverages on nodes' mobility and opportunistic contacts to enable proximity based file sharing. To this aim, we have integrated M2MShare in the Opportunistic Network Environment (ONE) simulator and tested its performance with realistic movement models in a map-based environment. This complex city-scale testbed includes a multitude of nodes representing real users engaged in their daily activities and movements, and allowed us to demonstrate the efficacy of M2MShare. 
\\

We evaluated M2MShare employing one-hop delegations against two other systems using different strategies:
\begin{itemize}
\item \textit{No\_delegation:} a system which does not employ delegations and file exchange is initiated only when a peer holding the requested data file is found in reach area of file requester.
\item \textit{Delegation\_to\_all:} a system employing delegations but instead employs the trivial technique where missing tasks are delegated to each encountered peer.
\end{itemize}
Our analysis proved that the use of delegations is more efficient than the strategy which does not employ them, due to the number of servant peers involved in the file-search activity. This number is limited using as heuristic the history of met nodes and delegating tasks only to frequently encountered peers. We proved that this strategy allows M2MShare to be more efficient than the strategy in which tasks are indiscriminately delegated to all met peers, with respect to the number of peers involved and to the redundancy introduced in the entire network.
\\

We improved M2MShare by adding support for multi-hop delegations and we proved how this increases the search area for a file, compared to the single-hop delegations strategy introduced by M2MShare.
\\

We compared M2MShare file division strategy with two other division strategies:
\begin{itemize}
\item iM: a strategy which requests at each file server the entire file, always starting from its first byte;
\item rM: a strategy that randomly chooses the initial download point in the file request.
\end{itemize}
and we proved that our file division strategy has the least redundancy during data transfer, especially considering big-sized files.
\\

Some of the previous results will be presented with a paper we wrote at Wireless Days Conference 2011 \footnote{Wireless Days 2011 \href{http://www.wireless-days.org/}{http://www.wireless-days.org/}} in Niagara Falls, Ontario, Canada. The full text of the paper is included in Appendix \ref{appendice-paper}.
Summarizing, our contributions are the following:
\begin{itemize}
\item implementing M2MShare and evaluating its behaviour using a simulator (the ONE simulator) able to emulate nodes movement in a realistic urban scenario;
\item evaluating the efficiency of the new paradigm created by M2MShare of use P2P solutions that matches file sharing with mobile users, allowing them to exchange files with each other;
\item evaluating the efficiency of using task delegations to dynamically establish forward routes along the destination path in the network;
\item evaluating the efficiency of the new file division strategy introduced with M2MShare; 
\item enhancing M2MShare adding support to multi-hop delegations in order to further increase the search area for a single node to other disconnected overlay networks;
\item enhancing the ONE simulator adding some new features to it and then making them available to the simulator users community\footnote{The ONE simulator user contributions \href{http://www.netlab.tkk.fi/tutkimus/dtn/theone/\#contrib}{http://www.netlab.tkk.fi/tutkimus/dtn/theone/\#contrib}}.
\end{itemize}

\pagebreak
\section{Future Work}
There are other analysis that could be done to evaluate M2MShare efficiency, with respect to several factors. Some of them might be:
%Some potential future developments to further improve and extend the functionalities provided by the software might be:
\begin{itemize}
\item Wider scenario: we emulated M2MShare behaviour and evaluated its efficiency in a city-scale scenario, adopting the Working Day Movement Model to emulate users daily activities. An interesting future work could be to evaluate M2MShare performance in a wider-scale environment, say a scenario composed by multiple cities and emulating commuters moving from one city to the other during their daily activities.
\item Power consumption analysis: one characteristic of mobile devices is the limited power. One future analysis could be to evaluate the impact of delegations in terms of energy consumption. Several models for  energy consumption analysis for Bluetooth are available and could be adapted and implemented in the ONE simulator.
\item Search phase emulation: we emulated only the second part of the protocol, the one in which a node looks for a specific file and asks to other peers for it, using delegations. A possible future analysis would be to evaluate the efficiency of the Search Module of M2MShare in indexing and returning results for a user query, using delegations even for this activity.
\end{itemize}

\newpage
\section{Acknowledgements}
First of all, I have to thank my parents, Ganni and Gabriela, for the support they provided me through my entire life, for all their love and encouragement and for believe in me in every circumstance.\\

I must express my gratitude towards my advisor, professor Claudio Enrico Palazzi, for his supervision, guidance and advices in writing this thesis. \\

My thanks to my co-advisor, Armir Bujari, for his precious help and willingness throughout the work which lead to the realization of this thesis.\\

Special thanks to Silvia, for her persistent confidence in me, for motivating me and for always pushing me to do my best.\\

Thanks to my friend Yuri, who helped me during all these years we lived together as roommates.\\

My thanks to Alberto, Luca and everyone that worked and studied with me, friends before colleagues.\\

The support and encouragement of many other people has been indispensable during the last years, so I would show my gratitude to my friends, to everyone who has given me something, to everyone that has been with me, and to everyone I have not mentioned here.\\

\begin{flushright}
\textit{Daniele}
\end{flushright}

 
% ---------------------------------------------------------------------------
%: ----------------------- end of thesis sub-document ------------------------
% ---------------------------------------------------------------------------

