% this file is called up by thesis.tex
% content in this file will be fed into the main document

%: ----------------------- name of chapter  -------------------------
\chapter{Introduction}\label{introduzione} % top level followed by section, subsection


%: ----------------------- paths to graphics ------------------------

% change according to folder and file names
%\graphicspath{{2-Consorzi/images/}}


%: ----------------------- contents from here ------------------------
Since the origins of civilization, mankind has had the need to share informations between places very distant from each other. Communication means have evolved enormously through the centuries from the first letters carried by runners. Riding couriers moved mails through greater distances allowing at the same time a faster communication between far cities. The methods of communication kept changing and in the 1840's the electric telegraph was invented and messages were sent much quicker through the first invisible means of communication. Just twenty years later, the first transoceanic submarine cable had been laid, and it has been possible to instantaneously send messages between America and Europe. Next, the telephone was invented and people began to talk from great distances as if they were sitting in the same room. Finally, with the rise of the Internet a new world-scale communication became possible.
\\

In the last years the access to the internet has become available for a large number of people, using many different types of devices to connect to the network. Fifteen years ago internet access was a prerogative of wired connected computers and cellular phones were devices only able to make phone calls and send/receive text messages. Today, smartphones are becoming very popular and these devices evolved from simple mobile phones to devices able to browse into the internet, read emails and watching video streaming from the net.
\\

This new request for connectivity issued new problems, as new connections frontiers became available
: while the network's core is highly connected and well suited for routing via
conventional routing algorithms, the network‘s expanding frontiers have infrastructure
that suffers from intermittent connectivity and changes in topology that can be difficult
or impossible to predict.




parlare a voce come essere presenti di persona
i telefoni sono diventati mobili garantendo la possibilità di parlarsi a distanza ovunque si sia, anche in mobilità
i telefoni cellulari si sono poi evoluti, dai primi modelli che permettevano solo telefonate voce e al massimo messaggi di testo, fino ad apparecchi che permettono delle performances fino a qualche anno fa riservate a del personal computer.
dal navigare in internet, controllare le email, ascoltare musica in streaming
tutte queste nuove posibilità hanno però sollevato dei nuovi requisiti di connettività in situazioni difficili da coprire tramite le soluzioni abitualmente utilizzate in ambienti cablati, come è la struttura classica di internet
elevata mobilità e limited power



\section{Contributions}
In this project, we evaluate the performance of M2MShare, a protocol which implements DTN techniques in mobile phones world, in order to enable a peer-to-peer file sharing system between peers in the network using their mobility and opportunistic contacts among them. To do so we implement the protocol in the Oppostunistic Network Environment (ONE) simulator. This simulator is able to emulate human movement adopting several movement models in a map-based environment. To our judgement, our contributions are the following:
\begin{itemize}
\item implementing M2MShare and evaluating its behaviour using a simulator (the ONE simulator) able to emulate nodes movement in a realistic urban scenario.
\item evaluating the efficiency of the new paradigm created by M2MShare of use P2P solutions that matches file sharing with mobile users, allowing them to exchange files with each other.
\item evaluating the efficiency of using task delegations to dynamically establish forward routes along the destination path in the network.
\item evaluating the efficiency of the new file division strategy introduced with M2MShare 
\item enhancing M2MShare adding support to multi-hop delegations in order to further increase the search area for a single node to other disconnected overlay networks.
\item enhancing the ONE simulator adding some features to it and then making them available to the simulator users community.
\end{itemize}

\section{Document outline}
The remainder of this document is organized with the following chapters:
\begin{enumerate}
\setcounter{enumi}{1}
\item \textbf{Background:} in this chapter we describe some must-know technology, useful to the reader before to read the following chapters. We describe Delay Tolerant Networks (DTN) and Peer-to-Peer (P2P) systems.
\item \textbf{M2MShare:} this chapter describes M2MShare, the protocol object of the study and simulations of this thesis.
\item \textbf{Movement models:} in this chapter we give an overview about movement models used to simulate people movements. We describe the characteristics of several movement models, comparing them, and finally we describe the model we use for our simulations.
\item \textbf{The ONE simulator:} this chapter presents the ONE (Opportunistic Network Environment), the simulator we used for our evaluation of M2MShare.
\item \textbf{Implementation:} in this chapter we describe our implementation of M2MShare into the ONE and the improvements we did to the simulator to completely emulate the protocol behaviour.
\item \textbf{Simulations and results:} in this chapter we present the simulations we did. For each simulation we describe the settings, the protocol aspect evaluated, and finally we give the related results with a graphical representation.
\item \textbf{Conclusions:}  this chapter gives the concluding remarks, and proposes some related future works and enhancements. %It also contains the due acknowledgements.

\end{enumerate}
 
% ---------------------------------------------------------------------------
%: ----------------------- end of thesis sub-document ------------------------
% ---------------------------------------------------------------------------

